Red Wine Analysis by Darragh Mac Ardle
========================================================



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(reshape2)
library(GGally)
library(gridExtra)
library(dplyr)
library(memisc)
```

```{r echo=FALSE, Load_the_Data}
wine <- read.csv("wineQualityReds.csv")
```

<br>
<br>
  
### Dataset introduction

1599 samples of red Portuguese "Vinho Verde" wines.
  
Description of attributes:

1. fixed acidity: most acids involved with wine or fixed or nonvolatile (do 
not evaporate readily)

2. volatile acidity: the amount of acetic acid in wine, which at too high of
levels can lead to an unpleasant, vinegar taste

3. citric acid: found in small quantities, citric acid can add 'freshness' 
and flavor to wines

4. residual sugar: the amount of sugar remaining after fermentation stops, 
it's rare to find wines with less than 1 gram/liter and wines with 
greater than 45 grams/liter are considered sweet

5. chlorides: the amount of salt in the wine

6. free sulfur dioxide: the free form of SO2 exists in equilibrium between 
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents
microbial growth and the oxidation of wine

7. total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO2 is mostly undetectable in wine, but at free SO2 
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of 
wine

8. density: the density of water is close to that of water depending on the 
percent alcohol and sugar content

9. pH: describes how acidic or basic a wine is on a scale from 0 (very 
acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

10. sulphates: a wine additive which can contribute to sulfur dioxide gas 
(S02) levels, wich acts as an antimicrobial and antioxidant

11. alcohol: the percent alcohol content of the wine
   
<br>
  
Output variable (based on sensory data): 

12. quality (score between 0 and 10). Median of at least 3 evaluations made 
    by wine experts.
       
<br>
  

Units:

1. Fixed acidity (tartaric acid - g / dm^3)

2. volatile acidity (acetic acid - g / dm^3)

3. citric acid (g / dm^3)

4. residual sugar (g / dm^3)

5. chlorides (sodium chloride - g / dm^3)

6. free sulfur dioxide (mg / dm^3)

7. total sulfur dioxide (mg / dm^3)

8. density (g / cm^3)

9. pH

10. sulphates (potassium sulphate - g / dm3)

11. alcohol (% by volume)

12. quality (score between 0 and 10)

<br>

```{r echo=FALSE}
str(wine)
```

Checking the structure of the dataset there's also a row counter variable "X" so 
there's 13 columns in the dataframe.

<br>
<br>
  

# Univariate Plots Section

### Quality

```{r echo=FALSE, Univariate_Plots, message=FALSE, warning=FALSE}
ggplot(data = wine, aes(x = factor(quality))) +
  geom_bar()
```

- 1 to 10 scale actual values are only between 3 and 8.
- Peak at 5 and 6. Makes sense as mean on a 1 to 10 scale would be 5.5
- Looks like there may be more wines above 5.5 than below.

This graph looks normally distributed, with a little positive skew as it looks 
like there's a longer tail on the right of the graph. This makes sense as the 
quality scores are the median of at least 3 expert reviews. From the cental 
limit therom we can should expect to see a normal distribution, even if all the
individual scores plotted a different distribution.


Creating a box plot to investigate further as I want to see the quartile ranges 
better. Overlaid a points plot get a better idea of the densities.

```{r echo=FALSE}
ggplot(data = wine, aes(x = "", y = quality)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.25, height = 0.25, color = 'red')
```

- Wines scoring 3 and 8 are identified as outliers.
- The median line can't be seen as it's at 6, on top of the 3rd quartile line.

<br>
<br>
  

### Alcohol

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = alcohol)) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks = seq(8, 15, by = 0.5))
```

- Positively skewed distribution.
- Big peak at 9.5% and some other, smaller peaks at 10.5 and 11%. This is likely
from a tendency to make alcohol levels to round numbers for marketing.

<br>
<br>
  
### Density

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = density)) +
  geom_histogram(binwidth = 0.0005)
```

- Normally distributed
- Very small range.

<br>
<br>
  
  
### Acidity related variables

This is fixed.acidity, volatile.acidity, citric.acid and pH


```{r echo=FALSE}
acid1 <- ggplot(data = wine,
                aes(x = fixed.acidity)) +
  geom_histogram(binwidth = 0.5)

acid2 <- ggplot(data = wine,
                aes(x = volatile.acidity)) +
  geom_histogram(binwidth = 0.02) +
  scale_x_continuous(breaks = seq(0, 1.6, by = 0.1))

acid3 <- ggplot(data = wine,
                aes(x = citric.acid)) +
  geom_histogram(binwidth = 0.02) +
  scale_x_continuous()

acid4 <- ggplot(data = wine,
                aes(x = pH)) +
  geom_histogram(binwidth = 0.02) +
  scale_x_continuous(breaks = seq(0, 4, by = 0.1))

grid.arrange(acid1, acid2, acid3, acid4, ncol = 1)
```

- fixed.acidity, volatile.acidity and citric.acid all positively skewed.
- volatile acidity has a bimodal distribution with peaks at roughly 0.42 and 
0.58.
- citric.acid has an spike at just below 0.5.
- pH is normally distributed

I find it surprising that pH is normally distributed given the other acid 
variables are all skewed to lower levels. I expected pH would have been skewed
towards more neutral (towards 7) with a long tail to lower pH values.

To investigate this further I'm creating a new variable that sums the 3 acid 
variables. They measure tartaric, acetic and citric acids and all have the same
units of g/litre so can be added together without modification.

This will hopefully give a rough guide to total acidity but may be flawed if 
the three acids have very different strengths.

```{r}
wine$total.acids <- with(wine, fixed.acidity + volatile.acidity + citric.acid)
```

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = total.acids)) +
  geom_histogram(binwidth = 0.2)
```

- Distibution is still a bit positively skewed.

<br>
<br>
  

### Sulphate and Sulpher Variables

```{r echo=FALSE}
sul1 <- ggplot(data = wine,
            aes(x = sulphates)) +
     geom_histogram(binwidth = 0.05)

sul2 <- ggplot(data = wine,
            aes(x = free.sulfur.dioxide)) +
     geom_histogram(binwidth = 2)

sul3 <- ggplot(data = wine,
            aes(x = total.sulfur.dioxide)) +
     geom_histogram(binwidth = 5)

grid.arrange(sul1, sul2, sul3)
```

- All 3 are skewed to the right.

Total.sulfur.dioxide is the very skewed so may benefit from a log transformation

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
            aes(x = total.sulfur.dioxide)) +
  geom_histogram() +
  scale_x_log10()
```

This gives a much more normal distribution which may be useful in later
analysis.

<br>
<br>
  

### Salt and Sugar (chlorides and residual.sugar variables)

Looking at chlorides (which is the salt content) and residual.sugar variables
together as these are two things that are normally labelled as bad for your 
health.

```{r echo=FALSE}
p1 = ggplot(data = wine,
            aes(x = chlorides)) +
     geom_histogram(binwidth = 0.01)

p2 = ggplot(data = wine,
            aes(x = residual.sugar)) +
     geom_histogram(binwidth = 0.5) +
     scale_x_continuous(breaks = seq(0, 16, by = 1))

grid.arrange(p1, p2, ncol = 2)
```


- similar distributions: big peaks low in the range and then a few much higher 
outliers. 

In the bivariate analysis section I'll investigate if it's the same wines that 
have both high salt and sugar or if it's more random.

<br>
<br>
  
  
  
# Univariate Analysis
  
### What is the structure of your dataset?
1599 observations of 11 input variables, which are all numeric, and one output
variable for quality which is between 1 and 10.

The first column of the dataset is an ID variable that doesn't contain any 
information past the row number.

  
### What is/are the main feature(s) of interest in your dataset?
The main feature of interest is quality. I want to look at what other variables
influence this and if a predictive model can be made from them.

  
### What other features in the dataset do you think will help support your \
### investigation into your feature(s) of interest?

I expect alcohol to have an effect on quality but it's hard to know intuitively 
how the other variables will affect it. Correclations can be measured to gauge 
which are the most important though.

  
### Did you create any new variables from existing variables in the dataset?

Added total.acids to help investigate why pH is normally distibuted while the 
3 different acids measured are all skewed to low levels. It will be interesting
to look at the relationship of this new variable to pH in the bivariate section.

  
### Of the features you investigated, were there any unusual distributions? 
### Did you perform any operations on the data to tidy, adjust, or change the 
### form of the data? If so, why did you do this?

Citric.acid was unusual having a high spike at just below 0.5 and being fairly
flat for most of it's distribution.

I was surprised to see that pH and density had normal distributions given that 
the variables that I expected to inflence them (fixed.acidity, volatile acidity
and citric acid for pH and alcohol and residual.sugar for density) all had 
skewed distributions.


<br>
<br>
  


# Bivariate Plots Section

Creating a pairs matrix to get a quick idea of the variables and their 
correlations.

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1836)
ggpairs(wine)
```

It's hard to see the correlations values on plot when it's not blown up to full
screen so remaking a plot for just the correlations:

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggcorr(wine, label = TRUE, layout.exp = 2, hjust = 0.9)
```


I see that alcohol and then volatile.acidity have the strongest correlations to 
quality so I shall start by looking at those. 

The two correlations are:

```{r}
cor.test(wine$quality, wine$alcohol)
cor.test(wine$quality, wine$volatile.acidity)
```
  
<br>
<br>
  
  
### Quality Vs Alcohol

In the graph jitter has been added horizontally but not vertically.

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = quality, y = alcohol)) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.25) +
  scale_x_continuous(breaks = seq(3, 8, by = 1)) +
  scale_y_continuous(breaks = seq(8, 15, by = 1)) +
  stat_smooth(method = "lm")
```

- The linear regression line shows a trend between increasing alcohol and 
increasing quality.
- The lowest alcohol content for wine with quality=8 is 9.5%. This looks to be 
above the median of the quality=5 wines. A boxplot chart may illustrate this 
better.
- The highest alcohol content off the quality=3 wines is 11%. This looks close 
median of the quality=6 wines and below that of the quality=7 wines. Again a 
boxplot would better visualise this.
- There seems to be an odd lower cutoff of alcohol point for the quality= 4, 5 
and 6 wines of 9%. There's a couple of outliers below this but it seems odd they
stop so suddenly.
- There looks to be a gap between wines at or near 13.5% and 14% percent. As 
there's a few wines on exactly 14% maybe there's a marketing reason makers aim 
for this.
- There's an outlier at nearly 15% alcohol for quality=5 and quality=4 also has 
a high indvidual outlier at just over 13%.


Remaking the graph as a boxplot.

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = factor(quality), y = alcohol)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(8, 15, by = 1))
```

- There's a lot of outliers in quality=5
- The change in median alcohol content is much clearer here than the linear 
regression line showed on the scatterplot. For quality=3, 4, 5 there's not much 
change. However from quality=5 to 8 there's a noticable step each time.
- The interquartile ranges for quality = 4, 6, 7, 8 are similar. It's odd that 
quality=5 has such a small range with so many outliers.

<br>
<br>
  

### Quality Vs volatile.acidity

Looking at volatile.acidity with quality with width but no height jitter.
```{r echo=FALSE}
ggplot(data = wine,
       aes(x = quality, y = volatile.acidity)) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.1) +
  stat_smooth(method = "lm") +
  scale_y_continuous(breaks = seq(0, 1.6, by = 0.2), limits = c(0, 1.6))
```

- A negative correlation between volatile.acidity and quality can be seen, 
decreasing volatile acidity with increasing quality.
- volatile.acidity values seem to bottom out at 0.2 for quality=5, 6 and 7.
- The ranges of acidity decease with increasing quality.

Creating a box plot of the same thing.

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = factor(quality), y = volatile.acidity)) +
  geom_boxplot() +
  scale_y_continuous()
```

- Shows a fairly linear decreasing trend between the medians from quality = 3 
to 7. Then 7 and 8 are similar.
- Like the quality vs alcohol graph again shows a lot of outliers for quality=5.

<br>
<br>
  

### Alcohol Vs volatile.acidity

As these two variables have the highest correlations with quality it is worth 
seeing if they have a correlation with each other. If they have a strong 
correlation to each other that would imply they're explaining the same part of 
quality's variation. It would be more interesting if they realatively 
independent as then they can be used to explain more of variation in quality 
and would allow for better prediction of it.

Checking the correlation first
```{r echo=FALSE}
cor.test(wine$alcohol, wine$volatile.acidity)
```

A correlation of -0.2 isn't very big. The scatterplot below illustrates this 
with a linear regression line with a small negative slope. That is to say that 
higher alcohol levels have generally lower acidity levels.

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = alcohol, y = volatile.acidity)) +
  geom_jitter(alpha = 0.2) +
  stat_smooth(method="lm")
```
<br>
<br>
  

### Other Acidity Variables

Using ggpairs on just the acidity and quality variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#pass only acidity and quality variables to ggpairs
ggpairs(wine[, c("fixed.acidity", "volatile.acidity", "citric.acid", 
                 "pH", "total.acids", "quality")])
```

The very linear scatterplot between fixed.acidity and total.acids stands out 
and highlights something I hadn't realised when I made the total.acids variable.
The range of fixed.acidity is far greater than volatile.acidity or citric acid.
Due to this total.acids is nearly equal to fixed.acidity. Their correlation is 
also nearly perfect at 0.996 which futher illustrates this point.

This makes total.acids irrelevant as fixed.acidity can be used in its place.

What would may be more useful is if total.acids was change to sum of normalised
acids. The 3 acid variables divided by their ranges.

```{r}
#normalise each variable by dividing by its range (max - min)
wine$total.acids <- with(wine, 
                         fixed.acidity / 
                           (max(fixed.acidity ) - min(fixed.acidity))
                         +
                           volatile.acidity / 
                           (max(volatile.acidity) - min(volatile.acidity)) 
                         +
                           citric.acid / 
                           (max(citric.acid) - min(citric.acid)))

#```{r echo=FALSE, message=FALSE, warning=FALSE}
ggpairs(wine[, c("fixed.acidity", "volatile.acidity", "citric.acid", 
                 "pH", "total.acids", "quality")])
```

total.acid is now a more interesting variable. It shows a high correlation with
both fixed.acidity and citric.acid but not volatile.acidity. It's interesting
that volatile.acidity doesn't move with the other two.

Unfortunately total.acid's correlation to pH is lower than fixed.acidity and its
correlation to quality is lower than the 4 other acid variables so I'm not sure
if I'll use it any further.

Below the relationship between fixed.acidity and pH is shown. The relationship 
betweent the two isn't great and doesn't fit a curving or linear line that well.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = pH, y = fixed.acidity)) +
  geom_jitter(alpha = 0.2) +
  stat_smooth(method = "lm", color = 'red') +
  stat_smooth(method = "loess")
```

Citric acid has a small relationship with quality and may be useful when 
modelling it.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = quality, y = citric.acid)) +
  geom_jitter(alpha = 0.2, height = 0, width = 0.25) 
```

- There's an odd gap in quality=7 and 8 values of citric acid around 0.25. I 
don't know what could be the cause of this.

<br>
<br>
  

### Sulphate and Sulpher Variables

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggpairs(wine[, c("free.sulfur.dioxide", "total.sulfur.dioxide", "sulphates", 
                 "quality")])
```

- Although still small sulphates has the higest correlation to quality
- Only total and free.sulfur.dioxide have a high correlation to each other here.

The difference between free and total sulfur dioxide is "bound" sulfur dioxide.
I wonder if a variable for this might yield any more information.

```{r echo=FALSE, message=FALSE, warning=FALSE}
wine$bound.sulfur.dioxide <- with(wine, total.sulfur.dioxide - 
                                    free.sulfur.dioxide)

ggpairs(wine[, c("free.sulfur.dioxide", "total.sulfur.dioxide", 
                 "bound.sulfur.dioxide", "sulphates", "quality")])
```

The bound.sulfur.dioxide has a higher correlation with quality and an even 
lower correlation with sulphates. Although bound SO2 and sulphates only have low
correlations with quality, the fact they don't correlate well with each other 
suggests they explain different parts to the variation in quality. This will be 
useful when making a model for quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = quality, y = bound.sulfur.dioxide)) +
  geom_jitter(alpha = 0.1, height = 0, width = 0.25) +
  stat_smooth(method = 'loess')
```

- An hump at quality=5 and 6 with lower values for 7 and 8.
- Two very high outliers at 7.
- The relatively flat best fit line show there's only a small correlation here.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = quality, y = sulphates)) +
  geom_jitter(alpha = 0.1, height = 0, width = 0.25) +
  stat_smooth(method = 'loess')
```

- Again only a small correlation visible.

<br>
<br>
  

### Salt Vs Sugar

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = chlorides, y = residual.sugar)) +
  geom_point(alpha = 0.2) +
  geom_smooth()
```

- Points are very clustered bottom right of the graph.
- High salt (chlorides) and high sugar levels seem to be exclusive. High salt 
levels only happen at low residual.sugar levels and vice versa.
- geom_smooth() gives what looks like a very overfitted line. There doesn't 
appear to be much relationship.

Checking the correlation to quantify the relationship.
```{r}
cor.test(wine$chlorides, wine$residual.sugar)
```

At 0.05 there's very little correlation between these two parameters.


<br>
<br>


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the 
### investigation. How did the feature(s) of interest vary with other features 
### in the dataset?

Alcohol and volatile.acidity had the strongest correlations to quality. 
Alcohol's relationship was positive and volatile.acidity's negative.
  
  
### Did you observe any interesting relationships between the other features 
### (not the main feature(s) of interest)?

Bound.sulfur.dioxide and sulphates both have a small relationship with quality
but not with each other. This may make them useful for explaining different 
parts of quality's variation in a model.
  

### What was the strongest relationship you found?

I created a new variable called total.acid which was the sum of the three acid
variables. It had a near perfect correlation to fixed.acidity and that helped 
me realise an oversight. Fixed.acidity is much bigger in magnitude than the 
other two acidity variables so when I summed them it accounted for nearly all 
the variation.

Due to this I modified the variable to be a sum of the 3 other variables 
normalised.

Aside from this pH and fixed.acidity had a high relationship of 0.68 which makes 
sense as they are two difference measures of acidity.
  
  
<br>
<br>
<br>
  
  
# Multivariate Plots Section
  
### Quality, alcohol and volatile.acidity

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = alcohol, y = volatile.acidity)) +
  geom_jitter(aes(color = factor(wine$quality)), alpha = 0.4) +
  scale_color_brewer(palette="Set1")
```

- The quality=5 wines cluster to the right of the graph.
- The quality=6 wines have a much broader range of alcohol percentages.
- The quality=7 wines have a similar range of alcohol percentages to quality=6 
but generally lower volatile.aciditys.
- The quality=8 wines look skewed to higher alcohol values.
  
  
Recreating this graph as a facet grid shows these trends clearer:

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = alcohol, y = volatile.acidity, color = factor(quality))) +
  facet_grid(. ~ quality) +
  geom_jitter(alpha = 0.4) +
  scale_color_brewer(palette="Set1")
```

<br>
<br>
  
### Quality, Sulphates and Bound.sulfur.dioxide


```{r echo=FALSE}
ggplot(data = wine,
       aes(x = sulphates, y = bound.sulfur.dioxide)) +
  geom_point(aes(color = factor(wine$quality))) +
  scale_color_brewer(palette="Set1")
```

Some outliers make it hard to see what's going on in the bottom left so will 
zoom in on that.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = sulphates, y = bound.sulfur.dioxide)) +
  geom_point(aes(color = factor(wine$quality)), alpha =0.5) +
  scale_color_brewer(palette="Set1") +
  xlim(c(0.25, 1.4)) +
  ylim(c(0, 130))
```

- For quality=5, 6, 7, 8 the mean can be seen moving up in sulphates and down in
bound.sulfur.dioxide.

This can be better illustrated with summary points showing the mean values for
each quality level as shown below.


```{r echo=FALSE, message=FALSE, warning=FALSE}
#create a new data set for means by quality level
#uses the dplyr library
mean.by.quality <- wine %>%
  group_by(quality) %>% 
  summarize(bound.mean = mean(bound.sulfur.dioxide),
            sulphates.mean = mean(sulphates),
            quality.count = n())

ggplot(data = wine,
       aes(x = sulphates, y = bound.sulfur.dioxide)) +
  geom_point(aes(color = factor(wine$quality)), alpha =0.5) +
  scale_color_brewer(palette="Set1") +
  geom_point(data = mean.by.quality,
             aes(x = sulphates.mean, y = bound.mean,
                 fill = factor(quality),
                 size = quality.count), 
             color = 'black', 
             pch = 21) + #pch=21 is a shape that allows outline and fill to be
                         #colored seperately
  guides(fill=FALSE) + #don't show color legend again, already one shown. 
  scale_size_continuous(range = c(4, 13)) +
  scale_fill_brewer(palette="Set1") +
  geom_line(data = mean.by.quality,
             aes(x = sulphates.mean, y = bound.mean)) +
  xlim(c(0.25, 1.4)) +
  ylim(c(0, 130))
```

The line shows an interesting relationship with a peak in the middle. Seems like
high leves of bound.sulfur.dioxide and low levels of sulphates can be associated 
with mediocraty. 

<br>
<br>
  
### Quality, salt and sugar

```{r echo=FALSE}
ggplot(data = wine,
       aes(x = chlorides, y = residual.sugar)) +
  geom_point(aes(color = factor(wine$quality)), alpha = 0.5) +
  scale_color_brewer(palette="Set1")
```

- Quality=5 are the most common high salt (chloride) wines.
- Quality=6 are the most common high sugar wines.
- There are very few quality=7 or 8 wines high in either salt or sugar.

It's hard to see what's going on in the dense cluster in the bottom left of the 
graph so creating the same graph with shorter limits to see it better.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = chlorides, y = residual.sugar)) +
  geom_point(aes(color = factor(wine$quality)), alpha = 0.5) +
  xlim(0.03, 0.15) +
  ylim(1, 4) +
  scale_color_brewer(palette="Set1")
```

Looks pretty random so can't draw any insights.

<br>
<br>
  
### Building a Linear Model

Showing the r code below so that it's visible what variables are added to each
model.

```{r}
m1 <- lm(quality ~ alcohol, data = wine)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + sulphates)
m4 <- update(m3, ~ . + bound.sulfur.dioxide)

#add all the other variables to a 5th variable just to see if it helps
m5 <- update(m4, ~ . + fixed.acidity + citric.acid + residual.sugar + chlorides +
               free.sulfur.dioxide + total.sulfur.dioxide + density + pH)

#need to have memisc library installed and loaded for mtable to work.
mtable(m1, m2, m3, m4, m5, sdigits = 3)
```

After alcohol and volatile.acidity adding other variables has a limited effect 
in increasing R-squared.

However even the best R-squared value of 0.361 is not very high showing that 
this model is not very good at predicting a quality score from the variables.

<br>
<br>
<br>


# Multivariate Analysis
  
### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Alcohol and volatile.acidity still showed the strongest relationship with 
quality.


### Were there any interesting or surprising interactions between features?

Bound.sulfur.dioxide and sulphates have an ineresting non linear relationship 
with quality. Quality=5 having a much higher mean level of bound SO2.

Salt and sugar don't have much realationship with quality.


### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

A model was created but it's best R-squared value was only 0.361 so it was not 
very good. This may be because it is a linear model and the relationships 
between the variables and quality are not all linear, such as quality and bound 
SO2. Other modelling methods that can take into account non linear relationships
may give a better result.


------
<br>
<br>
<br>
  
# Final Plots and Summary


### Plot One

```{r echo=FALSE, Plot_One}
ggplot(data = wine,
       aes(x = alcohol, y = volatile.acidity, color = factor(quality))) +
  facet_grid(. ~ quality) +
  geom_jitter(alpha = 0.4) +
  scale_color_brewer(palette="Set1") +
  xlab('Alcohol (% by volume)') +
  ylab('Volatile Acidity (g/dm^3)') +
  ggtitle('Acidity, Alcohol and Quality') +
  guides(color=guide_legend(title="Quality"))
```

### Description One

From quality=5 to 8 it can be seen how the mean alcohol is increasing and the 
mean volatile acidity decreasing. These are the strongest correlating variables
to quality.
  
<br>
<br>

### Plot Two
```{r echo=FALSE, Plot_Two, messages=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = sulphates, y = bound.sulfur.dioxide)) +
  geom_point(aes(color = factor(wine$quality)), alpha =0.5) +
  scale_color_brewer(palette="Set1") +
  geom_point(data = mean.by.quality,
             aes(x = sulphates.mean, y = bound.mean,
                 fill = factor(quality),
                 size = quality.count), 
             color = 'black', 
             pch = 21) + #pch=21 is a shape that allows outline and fill to be
                         #colored seperately
  guides(fill=FALSE) + #don't show color legend again, already one shown. 
  scale_size_continuous(range = c(4, 13)) +
  scale_fill_brewer(palette="Set1") +
  geom_line(data = mean.by.quality,
             aes(x = sulphates.mean, y = bound.mean)) +
  xlim(c(0.25, 1.4)) +
  ylim(c(0, 130)) +
  xlab('Sulphates (g/dm3)') +
  ylab('Bound Sulfur Dioxide (mg/dm^3)') +
  ggtitle('Bound SO2, Sulphates and Quality') +
  guides(color=guide_legend(title="Quality")) +
  guides(size=guide_legend(title="Count at quality level"))
```

### Description Two

The line between the mean points shows an interesting relationship. Bound SO2 
and quality don't have a linear relationship. There's a peak at quality=5. This
shows that increasing bound SO2 is initally benificial but then a hindrence to 
reaching high quality levels. Whereas increasing sulphates has a more linear
relationship with higher levels correlating with higher quality scores.

<br>
<br>

### Plot Three
```{r echo=FALSE, Plot_Three, message=FALSE, warning=FALSE}
ggplot(data = wine,
       aes(x = chlorides, y = residual.sugar)) +
  geom_point(aes(color = factor(wine$quality)), alpha = 0.2) +
  scale_color_brewer(palette="Set1") +
  xlab('Salt (g/dm^3)') +
  ylab('Sugar (g/dm^3)') +
  ggtitle('Salt, Sugar and Quality') +
  guides(color=guide_legend(title="Quality"))
```

### Description Three

It is interesting that there are no wines that are high in both salt and sugar.
This was unexpected as it's common for other drinks to contain high levels of
both to improve their taste.

It can be seen that quality=5 are the most common high salt wine, qualit=6 are 
most common high sugar wines but the vast majority of wines are clustered with
both low salt and sugar. 


<br>
<br>


# Reflection

It was interesting that alcohol percentage had the strongest correlation to 
quality of the variables. The strenght of a wine is rarely discussed as key 
component of how good it will taste. The fact it had the strongest correlation
may point to a basic fact that connoisseurs see as beneath them: that people
often just like alcohol.

I stuggled to create an new variables that had a strong correlation to quality.
Bound.sulfur.dioxide was the best I did and it did have some limited use.

I was surprised that the acidity variables didn't have much correlation to the 
pH. It seemed natural to think that they woudl have been what was driving the 
pH.

I was also surprised that salt and sugar didn't have a correlation to quality or
each other as they are so commonly used to improve the taste in other drinks.

Future work could be done on investigating non linear modelling methods that 
could be used to make better predicitons of quality from the input variables.

